---
title: "Reproducable research project 1 2021"
author: "Guy Mathys"
date: "16-4-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Loading and preprocessing the data

## Read data
```{r ,echo = TRUE}
library(plyr)
library(ggplot2)
library(chron)

if (dir.exists ("C:/Users/Aureus/Documents/R/Data/DataScienceReproRes"))  {   
  setwd ("C:/Users/Aureus/Documents/R/Data/DataScienceReproRes") }

unzip("activity.zip")
data_raw <- read.csv("activity.csv")
data <- na.omit(data_raw)     

```
## mean total number of steps taken per day
```{r,echo = TRUE }

n_steps_total <- sum(data$steps)
n_steps_by_day <- aggregate(x= data$steps,
                            by= list(data$date),
                            FUN=sum)

hist(n_steps_by_day$x,
     main="Steps by day",
     xlab="steps",
     ylab = "days",
     col="blue")

```
Mean steps by day
```{r,echo = TRUE }

mean_1 <- mean(n_steps_by_day$x)
print(mean_1)

```
Median steps by day
```{r,echo = TRUE }
median_1 <- median(n_steps_by_day$x)
print(median_1)


```
# average daily activity pattern
## time series plot
```{r,echo = TRUE }
theme_set(theme_light())
n_steps_by_interval <- aggregate(x= data$steps,
                                 by= list(data$interval),
                                 FUN=mean)
colnames(n_steps_by_interval) <- c("interval", "steps")
ggplot(n_steps_by_interval, aes(x=interval)) +
  geom_line(aes(y=steps))

```
## The 5-minute interval that, on average, contains the maximum number of steps
```{r,echo = TRUE }
print ( n_steps_by_interval[which.max(n_steps_by_interval$steps),])

```
# Imputing missing values

total number of missing values in the dataset 
```{r,echo = TRUE }

print(nrow(data_raw) - nrow(data))

data_new <- data_raw

```
## new dataset

Strategy: The missing data (NA) is replaced by the mean of that interval over all days.
```{r,echo = TRUE }
range_1 <- 1: length(n_steps_by_interval$interval)
range_2 <- 1 : length(data_new$interval)

for (i1 in range_1)
{
  for(i2 in range_2)
  {
    if (n_steps_by_interval$interval[i1] == data_new$interval[i2])
    {
     if( is.na(data_new$steps[i2]))
      {
        data_new$steps[i2] <- n_steps_by_interval$steps[i1]
             }
      
    }
  }
}


n_steps_total_new <- sum(data_new$steps)
n_steps_by_day_new <- aggregate(x= data_new$steps,
                            by= list(data_new$date),
                            FUN=sum)

hist(n_steps_by_day_new$x,
     main="Steps by day",
     xlab="steps",
     ylab = "days",
     col="darkmagenta")


```
Mean steps by day
```{r,echo = TRUE }
mean_new <- mean(n_steps_by_day_new$x)
print(mean_new)
print ("difference")
print (mean_1 - mean_new)

```
Median steps by day
```{r,echo = TRUE }
median_new <- median(n_steps_by_day_new$x)
print ("difference")
print(median_1 - median_new)

```

# differences in activity patterns between weekdays and weekends
```{r,echo = TRUE }
#data_new$day <-  weekdays( as.Date(data_new$date))    

weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday','Friday','maandag','dinsdag','woensdag','donderdag','vrijdag')
data_new$day <- weekdays( as.Date(data_new$date))
data_new$day <- c('weekend', 'weekday')[(weekdays( as.Date(data_new$date)) %in% weekdays1)+1L]

week1 <- data_new$day == "weekend"
data_weekend <- data_new[week1,]
week2 <- data_new$day == "weekday"
data_week <- data_new[week2,]



```

## weekend
```{r,echo = TRUE }


  theme_set(theme_light())
  n_steps_by_interval <- aggregate(x= data_weekend$steps,
                                   by= list(data_weekend$interval),
                                   FUN=mean)
  colnames(n_steps_by_interval) <- c("interval", "steps")
  ggplot(n_steps_by_interval, aes(x=interval)) +
    geom_line(aes(y=steps))
  
 
```

### weekday
```{r,echo = TRUE } 
  n_steps_by_interval <- aggregate(x= data_week$steps,
                                   by= list(data_week$interval),
                                   FUN=mean)
  colnames(n_steps_by_interval) <- c("interval", "steps")
  ggplot(n_steps_by_interval, aes(x=interval)) +
    geom_line(aes(y=steps))
  
  
  

```






